---
output:
  pdf_document: default
  html_document:
    keep_md: no
---
```{r global_options, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```
# Executive Summary

Motor Trend, a magazine about the automobile industry, has analysed a data set of a collection of cars to explore the relationship between miles per gallon (MPG) and other variables. The magazine is particularly interested in the following two questions:

* "Is an automatic or manual transmission better for MPG"
* "Quantify the MPG difference between automatic and manual transmissions"

The analysis found the data supports the claim that cars with manual transmisison deliver better MPG than those with automatic.  The increase from automatic to manual transmission is 2.9 MPG.

# The Data

The data was extracted from the 1974 Motor Trends magazine and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models).  The list of variables can be found in the Appendix at Figure 1.

```{r}
library(UsingR)
data(mtcars)
library(ggplot2)
library(GGally)
```

# Exploratory data analysis

An understanding of the `am` variable means it should be a factor variable indicating the transmission type.

```{r}
# am is a factor variable (0 = automatic, 1 = manual)
mtcars$am = as.factor(ifelse(mtcars$am==0, "Automatic", "Manual"))
```

Figure 2 in the Appendix is a boxplot of the comparison of the mpg between manual and automatic transmission, showing that overall manual has higher mpg than automatic.  Figure 3 shows the relationship between all the variables in the dataset.  This assists with selection of model variables.

We perform a t-test where the null hypothesis is that there is no difference between automatic and manual transmissions.  The results in Figure 4 indicate the null hypothesis should be rejected (p value less than 5%), there is a statistical difference between automatic and manual transmissions on mpg.

```{r, results='hide'}
t.test(mtcars[mtcars$am == "Automatic",]$mpg, mtcars[mtcars$am == "Manual",]$mpg)
```

# Model selection

Multiple models will be created with mpg as the outcome and selected variables as predictors, the best fit will be identified.  Another model will be chosen by AIC in a stepwise algorithm.  The best fit model will be selected out of all of them.

```{r}
# create models with the most significantly influencing variables
fit0 <- lm(mpg ~ ., data = mtcars)
fit1 <- lm(mpg ~ am, data = mtcars)
fit2 <- lm(mpg ~ am + wt, data = mtcars)
fit3 <- lm(mpg ~ am + wt + hp, data = mtcars)
```

Using ANOVA, model 3 is identified as the best fit (see Figure 5).  Model 4 is created by using AIC in a stepwise algorithm to identify the best model from all available variables.

```{r}
fit4 <- step(fit0, trace=0)
```

We compare the summary of models 3 and 4 (see Figure 6 & 7) and select model 4 as best due to its slightly better R squared.  R squared indicates the percentage of variance explained by the model.

# Residual plots and diagnostics

Patterns in residual plots generally indicate some poor aspect of model fit.  Figure 8 contains the residual plots to demonstrate good model fit.  The **residual vs fitted** plot shows the points lie above and below the zero with no systematic patterns or large outlying observations.  The **scale-location** plot shows similar data, but the Y-axis is standardised residuals (rescaled so they have a mean of zero and a variance of one).  The red line in the plot should be relatively flat, which it is apart from the start and end.

The **normal Q-Q** plot indicates the residual to normal quantiles is good as they fall mostly along the line.  The **residuals vs leverage** plot shows the points are equally centred around zero, the red line is close to the horizontal dashed line and no points have a large Cook's distance (>0.5), which means the model is good.

# Conclusion

The coefficients for model 4 (Figure 7) indicate automatic transmission cars average 9.6 miles per gallon and manual cars deliver an additional 2.9 mpg.  The adjusted R-squared indicates an 83% explanation of the variance and the model's p-value is just less than the 5% cut off.  See Figure 6 & 7 for more detail.

# Appendix

Figure 1. The variables of the dataset are:

* `mpg` - Miles/(US) gallon
* `cyl` - Number of cylinders
* `disp` - Displacement (cu.in.)
* `hp` - Gross horsepower
* `drat` - Rear axle ratio
* `wt` - Weight (1000 lbs)
* `qsec` - 1/4 mile time
* `vs` - V/S
* `am` - Transmission (0 = automatic, 1 = manual)
* `gear` - Number of forward gears
* `carb` - Number of carburetors

```{r, echo=FALSE, cache=TRUE, fig.width = 12, fig.height = 8}
boxplot(mpg ~ am, data = mtcars, main = "Figure 2. MPG by Transmission Type", ylab = "MPG")

ggpairs(
 mtcars,
 mapping = ggplot2::aes(color = am),
 upper = list(continuous = wrap("density", alpha = 0.5), combo = "box"),
 lower = list(continuous = wrap("points", alpha = 0.3), combo = wrap("dot", alpha = 0.4)),
 title = "Figure 3. Pair Relationships"
)
```

Figure 4. T-test results

```{r, echo=FALSE}
t.test(mtcars[mtcars$am == "Automatic",]$mpg, mtcars[mtcars$am == "Manual",]$mpg)
```

Figure 5. ANOVA to identify best model (0-3)

```{r, echo=FALSE}
anova(fit1, fit2, fit3, fit0)
```

Figure 6. The best model identified by ANOVA

```{r, echo=FALSE}
summary(fit3)
```

Figure 7. The best model identified by AIC in a stepwise algorithm

```{r, echo=FALSE}
summary(fit4)
```

Figure 8. Residual plots

```{r, echo=FALSE, fig.width = 12, fig.height = 8}
par(mfrow=c(2,2))
plot(fit4)
```
