---
output:
  html_document:
    keep_md: no
---
```{r global_options, include = FALSE}
knitr::opts_chunk$set(fig.width = 12, fig.height = 8, fig.path = 'figure/',
                      echo = TRUE, warning = FALSE, message = FALSE)
```
# Introduction

Motor Trend, a magazine about the automobile industry, has looked at a data set of a collection of cars to explore the relationship between a set of variables and miles per gallon (MPG). The magazine is particularly interested in the following two questions:

* "Is an automatic or manual transmission better for MPG"
* "Quantify the MPG difference between automatic and manual transmissions"


# Executive Summary

The analysis identified data that supports the claim that cars with a manual transmisison deliver more (2.9mpg) than automatic cars.generally held belief that manual cars are provide more miles per gallon (mpg).


# The data

The data was extracted from the 1974 Motor Trends magazine and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models).

The variables of the dataset are:

* `mpg`   Miles/(US) gallon
* `cyl`	  Number of cylinders
* `disp`	Displacement (cu.in.)
* `hp`	  Gross horsepower
* `drat`	Rear axle ratio
* `wt`	  Weight (1000 lbs)
* `qsec`	1/4 mile time
* `vs`	  V/S
* `am`	  Transmission (0 = automatic, 1 = manual)
* `gear`	Number of forward gears
* `carb`	Number of carburetors

Load the data and libraries to process.

```{r}
library(UsingR)
data(mtcars)
library(ggplot2)
library(GGally)
```


# Exploratory data analysis

An understanding of the `am` variable means it should be a factor variable indicating the transmission type.

```{r}
# am is a factor variable (0 = automatic, 1 = manual)
mtcars$am = as.factor(ifelse(mtcars$am==0, "Automatic", "Manual"))
```

The plots for the exploratory data analysis can be found in the appendix.

Figure 1 is a boxplot of the comparison of the mpg between manual and automatic transmission, showing that over manual has higher mpg than automatic.

```{r}
boxplot(mpg ~ am, data = mtcars, main = "Fig 1. MPG by Transmission Type", ylab = "MPG")
```

Figure 2 shows the relationship between the variable in the dataset.  This assists with selection of model variables.

```{r, cache=TRUE}
pairs(~ ., data = mtcars, main = "Fig 2. Pair Relationships")

ggpairs(
 mtcars,
 mapping = ggplot2::aes(color = am),
 axisLabels = "internal",
 upper = "blank", #list(continuous = wrap("density", alpha = 0.5), combo = "box"),
 #lower = list(continuous = wrap("points", alpha = 0.3), combo = wrap("dot", alpha = 0.4)),
 title = "Cars"
)

ggpairs(
 mtcars,
 mapping = ggplot2::aes(color = am),
 upper = list(continuous = wrap("density", alpha = 0.5), combo = "box"),
 lower = list(continuous = wrap("points", alpha = 0.3), combo = wrap("dot", alpha = 0.4)),
 title = "Cars"
)
```

Perform a t test where the null hypothesis is that there is no difference between automatic and manual transmissions.

```{r, echo = FALSE}
t.test(mtcars[mtcars$am == "Automatic",]$mpg, mtcars[mtcars$am == "Manual",]$mpg)
```

In this case the null hypothesis should be rejected (p value less than 5%), there is a statistical difference between automatic and manual transmissions on mpg.  It would be worth while modeling the impact of transmission on mpg.


# Model selection

Multiple models will be created with mpg as the outcome and selected variables as predictors, the best fit will be identified.  Another model will be chosen by AIC in a stepwise algorithm.  The best fit model will be selected out of all of them.

```{r}
# create models with the most significantly influencing variables
fit0 <- lm(mpg ~ ., data = mtcars)
fit1 <- lm(mpg ~ am, data = mtcars)
fit2 <- lm(mpg ~ am + wt, data = mtcars)
fit3 <- lm(mpg ~ am + wt + hp, data = mtcars)
```

Use ANOVA to identify the best model.

```{r}
anova(fit1, fit2, fit3, fit0)
```

Using ANOVA model 3 is identified as best

```{r}
summary(fit3)
```

Use AIC in a stepwise algorithm to identify the best model from all the available variables.

```{r}
fit4 <- step(fit0, trace=0)
```

The best model includes `wt`, `qsec`, and `am`.

```{r}
summary(fit4)
```

Comparing model 3 and 4, model 4 is selected due to its slightly better R squared.  R squared indicates the percentage of variance explained by the model.


# Residual plots and diagnostics

Is the selected model good enough?  Patterns in residual plots generally indicate some poor aspect of model fit.

The residual vs fitted plot shows the points lie above and below the zero with no systematic patterns or large outlying observations.  The scale-location plot shows similar data, but the Y-axis is standardised residuals (rescaled so that they have a mean of zero and a variance of one).  The red line in the plot should be relatively flat, which it is apart from the start and end.

The normal Q-Q plot indicates the residual to normal quantiles is good as they fall mostly along the line.  The residuals vs leverage plot shows the points are equally centred around zero, the red line is close to the horizontal dashed line and no points have a large Cook's distance (>0.5), which means the model is good.

```{r}
par(mfrow=c(2,2))
#plot(fit0, main="fit0")
#plot(fit1, main="fit1")
#plot(fit2, main="fit2")
#plot(fit3, main="fit3")
plot(fit4)
```


# Conclusion

The coefficients for model 4 indicate automatic transmission cars average 9.6 miles per gallon and manual cars delivering an additional 2.9 mpg.  The adjusted R-squared indicates a 83% fit with the variance and the model's p-value is just less than the 5% cut off.


## Question

Take the mtcars data set and write up an analysis to answer their question using regression models and exploratory data analyses.

Your report must be:

* Written as a PDF printout of a compiled (using knitr) R markdown document.
* Brief. Roughly the equivalent of 2 pages or less for the main text. Supporting figures in an appendix can be included up to 5 total pages including the 2 for the main report. The appendix can only include figures.

Peer Grading

* Did the student interpret the coefficients correctly?
* Did the student do some exploratory data analyses?
* Did the student fit multiple models and detail their strategy for model selection?
* Did the student answer the questions of interest or detail why the question(s) is (are) not answerable?
* Did the student do a residual plot and some diagnostics?
* Did the student quantify the uncertainty in their conclusions and/or perform an inference correctly?
* Was the report brief (about 2 pages long) for the main body of the report and no longer than 5 with supporting appendix of figures?
* Did the report include an executive summary?
* Was the report done in Rmd (knitr)?

## Other stuff

```{r}
#Not sure whether to include vif().
library(car)
vif(fit4)
sqrt(vif(fit4))

#sort(cor(mtcars)[1,])

fit = lm(mpg ~ am, data = mtcars)
auto = mtcars[mtcars$am=="Automatic",]
man = mtcars[mtcars$am=="Manual",]

plot(mtcars$wt, mtcars$mpg, type = "n", frame = FALSE)
abline(lm(mtcars$mpg ~ mtcars$wt), lwd = 2)
abline(h = mean(auto$mpg), lwd = 3)
abline(h = mean(man$mpg), lwd = 3)
fit <- lm(mpg ~ wt + am, data = mtcars)
abline(coef(fit)[1], coef(fit)[2], lwd = 3)
abline(coef(fit)[1] + coef(fit)[3], coef(fit)[2], lwd = 3)
points(auto$mpg, pch = 21, col = "black", bg = "lightblue", cex = 2)
points(man$mpg, pch = 21, col = "black", bg = "salmon", cex = 2)

g = ggplot(mtcars, aes(x = wt, y = mpg))
g = g + xlab("Weight")
g = g + ylab("MPG")
g = g + geom_point(data = mtcars, aes(x = wt, y = mpg, color = am), size = 4)
g

pairs(~mpg+am+wt+hp,data=mtcars, main="Simple Scatterplot Matrix")
pairs(mtcars, panel = panel.smooth, main = "mtcars Dataset", col = 3 + (mtcars$am == "Automatic"))
boxplot(mpg ~ am, data = mtcars, main = "Transmission Type", ylab = "MPG")

#The gclus package provides options to rearrange the variables so that those with higher correlations are closer to the principal diagonal. It can also color code the cells to reflect the size of the correlations.
# Scatterplot Matrices from the glus Package
#install.packages("gclus")
library(gclus)
dta <- mtcars[c(1,3,5,6)] # get data
dta.r <- abs(cor(dta)) # get correlations
dta.col <- dmat.color(dta.r) # get colors
# reorder variables so those with highest correlation
# are closest to the diagonal
dta.o <- order.single(dta.r)
cpairs(dta, dta.o, panel.colors=dta.col, gap=.5,       
       main="Variables Ordered and Colored by Correlation" )

fit <- lm(mpg ~ hp, data = mtcars)
temp <- mtcars
temp$resid <- resid(fit)

g <- ggplot(temp, aes(x = am, y = resid))
g <- g + geom_hline(yintercept = 0, col="red")
g <- g + geom_point(alpha = .5, cex = 5)
g

#http://data.princeton.edu/R/linearModels.html
par(mfrow=c(2,2))
plot(fit)

plot(fitted(fit), residuals(fit), xlab = "Fitted Values", ylab = "Residuals")
abline(h=0, lty=2)
lines(smooth.spline(fitted(fit), residuals(fit)))
```
